\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-21 paper/paper.tex
Updates to paper.

\documentclass[10pt, preprint, nonatbib]{sigplanconf}
\usepackage{times,epsfig,endnotes,graphicx, verbatim}
\usepackage{datetime, url, pdfpages}

@@

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions - with innovations targeted at taking
advantage of growing flash speeds. Unfortunately their performance is hampered
by the need to maintain and replicate a write ahead log to guarantee
availability in the face of machine and storage failures. Cyclone offers a
solution to this problem by leveraging a small amount of directly attached
non-volatile memory (NVM) such as NVDIMMs to maintain a two level log - the upper
level in a small amount of NVM draining into a larger lower level log
maintained on flash. Replicating the smaller log kept in NVMs transforms the
replication problem into a packet switching problem - the leader sends the exact
same packet to all follower replicas and we solve the problem as such by using
the RAFT protocol implemented as a software multicast module on the Data Plane
Development Kit (DPDK). Second, we levergage the observation that key value store
interfaces are commutative for operations to different keys to scale our two
level log horizontally into a number of physical two level logs - with a hash of
the key being used to choose the right log. Each log is replicated by an
independent instance of RAFT, thereby leveraging the scalability afforded by
classic software packet switching when using multiple threads. Cyclone is able
to replicate millions of small updates a second using only commodity 10 gigabit
ethernet adapters and improves the performance of RocksDB - a popular persistent
key value store by orders of magnitude when compared to its own write ahead log
without replication.
\end{abstract}

\section{Introduction}
Persistent key value stores are an increasingly important component of
datacenter scale storage services. Key value stores such as
Rocksdb~\cite{rocksdb}, LevelDB~\cite{leveldb} and FloDB~\cite{flodb} represent
large amounts of effort on both the engineering and deployment fronts. These key
value stores represent sophisticated in-memory data structures built around Log
Structured Merge trees~\cite{lsmtree} and are heavily tuned to extract maximum
performance from present and future generations of flash solid state drives
(SSDs).

These key value stores however have a hitherto ignored an important component -
the write ahead log. A machine or storage failure leading to total loss or
temporary unavailability of data is unacceptable in many internet scale services
where high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write-ahead log that if replicated
and made durable for every appended update provides the necessary high
availability. Unfortunately the write ahead is a performance achilles heel for
these systems - rendering much of the work on improving performance of the LSM
component moot. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked Rocksdb shows the performance of
Rocksdb without the write ahead log. Both persisting the write ahead log without
replicating it - the line marked Rockdb/WAL as well simply replicating the log
three ways using the RAFT~\cite{raft} protocol over TCP/IP without persisting it
are at least an order of magnitude slower (observe the log scale on the
x-axis). It is therefore no suprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store researchers such as the FloDB~\cite{flodb} authors turn off the write
ahead log to be able to showcase benefits of sophisticated extensions to LSM
datastructures.

@@

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone provides the
order of magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the
server - such as in the form of NVDIMMs~\cite{farm} or directly attached
persistent memory~\cite{pmfs}.

Cyclone makes use of a limited amount of attached NVM using two simple
ideas. First it leverages the observation that with persistence being a first
class property of a subset of directly addressable DRAM, replicating the same
log entry from the leader to follower replicas is no different from multicast
using software packet switching techniques. It directly leverages this
observation to implement a version of RAFT the replicates the NVM log using
classic packet switching techniques such as separation of control and
dataplanes. Second, Cyclone uses a two level log structure where it drains
entries from the NVM log in IO friendly batches to a log place on a flash
drive. This allows Cyclone to use only a limited amount of NVM as current forms
of NVM such as NVDIMMs are more expensive than DRAM. Our experiments use an NVM
footprint of a mere 64 MB to achieve performance comparable to placing the
entire log in NVM. This allows Cyclone to be used in deployments where the key
value store in maintanied entirely in main memory with the log serving as the
source for recovery in the event of failure, rather than a multi-level LSM tree
structure.

We discuss the two level structure focusing on the packet switching components
in Section~\ref{sec:2level}. We then describe how we can horizontally scale this
basic idea across multiple cores by leveraging the commutativity of key value
store APIs across different keys in Section~\ref{sec:horizontal}. We then
demonstrate in the evaluation in Section~\ref{sec:evaluation} that Cyclone can
replicate millions of updates a second using only commodity 10 Gigabit ethernet
thereby closing the performance gap between key value store performance and
write ahead logging performance using only a small amount of directly attached
NVM. A discussion of related work follows before we conclude.

\section{Two Level Log}
\label{sec:2level}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone'e construction is predicated on the observation that
in the common case protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in persistent directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flash log. We periodically drain entries in batches from the
NVM log to the backing flashlog. An important and beneficial side effect of this
two level approach is that we need only limited amounts of NVM with the bulk of
the log residing on flash - an arrangement that maps well to the relative cost
of the two types of memory. We begin by describing the NVM log and how we
implement the RAFT consensus protocol in the packet switching paradigm around
it. We then describe the flashlog and how we drain entries in batches from the
NVM level to the flash level.

\subsection{NVM Log}
The NVM log is maintained as a circular log of fixed sized pointers (a pointer
log) to memory buffers - each of which holds a client request. This arrangement
is shown in Figure~\ref{fig:nvm_log} that also illustrates the advantage of
such an arragement. Adding a level of indirection in the NVM log allows us to
isolate the circular log being manipulated by the RAFT protocol from the
problems of variable sized client requests that can be scattered across
non-volatile memory depending on the internals of the memory allocator of the
packets switching library we use. Another advantage is that it makes recovery
from NVM easy - appends to the circular pointer log are atomic and we use the
pointer log to recover allocator state i.e. what pieces of NVM are currently in
use by the log.

@@

We now describe how we implement the RAFT consensus protocol to replicate the
NVM log across a set of replicas. The software packet switching idea we apply is
to separate the control plane from the dataplane using the Data Plane
Development Kit (DPDK~\cite{dpdk}) library for direct userspace access to the
network interface.

\subsubsection{Dataplane Separation}
We use RAFT~\cite{raft} as the underlying protocol for log replication. RAFT
replicates a log across a set of replicas. Its operations are leader-based and
RAFT incorporates a leader election algorithm that elects a leader with the most
upto date log. The leader receives log entries from clients, appends them to the
persistent log (NVM log in this case) and multicasts the new entry to follower
replicas. On a response from enough follower replicas (constituting a majority
quorum) the leader replica can then execute the command in the log entry. In
our case this involves accessing the key value store and sending the response
back to the client.

We implement the common case flow in RAFT as a software packet switch using
DPDK. The dataplane now encompasses the NIC and the CPU caches with the NIC
placing packet data directly in the CPU cache via DDIO~\cite{ddio} - a standard
feature of most NICs today. The dataplane for the RAFT protocol is therefore
separated from the control plane. The control plane now runs in a thread pinned
to a dedicated CPU core and takes care of checking adherence to the protocol:
checking for correct term, log index and that the replica receiving the packet
from the client is indeed the leader. In addition, the control plane is
responsible for non common-case activities such as dealing with timeout and
leader elections.

The pseudocode in Figure~\ref{fig:control_plane} describes the control plane in
Cyclone organized as event handlers triggered on receiving a packet. We focus on
two events for illustration (and brevity): the event where a request is received
from a client and the event where a leader election related message is
received. A key concern when we built Cyclone was adhering as far as possible to
the software packet switching principle of separating the control plane from the
dataplane. This means that the CPU should avoid examining every byte of the
packet as far as possible.

If an election message is received all the code related to handling
the election related message is run on the CPU core and all contents of the
packet are examined to decide what to do in accordance with the RAFT protocol
state machine. This is an exception to the software packet switching principle
but one that we found acceptable since leader elections are rare compared to the
common case of replicating client packets.

The sequence of operations when a packet is received from the client however
needs careful design for efficient software packet switching. To illustrate how
this is achieved, Figure~\ref{fig:packet_layout} shows how Cyclone manipulates
packet layouts across the two steps of prepending a RAFT header and transmitting
to follower replicas. DPDK describes packets using an ``mbuf'' data
structure. Roughly speaking, an ``mbuf'' consists of a flat array of bytes
actually containing the byte and a fixed size piece of external metadata that
describes various aspects of the packet, most crucially a pointer to the start
and end of the packet in the byte array. DPDK's userspace drivers receive
packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We strip off the existing
network headers in the packet and prepend RAFT related information specific to
each log entry in the headroom by shifting the start pointer appropriately -
these operations are standard enough for software packet switches that DPDK
provides convenient library calls to do it. For the final step, we need to
prepare the packet for transmission to the various follower replicas. To do this
we prepare a different packet containing an ethernet header for the targeted
replica and ``chain'' the data packet to each of these headers. Each header is
then separately handed off to the driver for transmission via the NIC, carrying
the data packet with it by association.

\begin{figure}
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }

    event_receive_election_msg()
    {
      Handle_election_msg()
    }
\end{verbatim}
\caption{Cyclone Control Plane Fragment}
\label{fig:control_plane}
\end{figure}

@@

We note that by discrding stream oriented protocols the network layer is not
longer reliable or ordered. This is not a problem for RAFT that is designed to
tolerate network failures resulting in drops or reordering of network
packets. However this rules out consensus protocols such as Zookeeper Atomic
Broadcast (ZAB~\cite{zab}) that depend on a reliable stream-oriented network
layer to function correctly.

@@

We note there that batching in Cyclone does not involve a latency-throughput
tradeoff like in many other systems~\cite{ix-dataplane}. The batch receive call
we use in DPDK returns immediately with whatever number of packets is available,
including zero. We always flush the transmit buffer after every call to DPDK to
transmit packets to replicas. Therefore, we never tradeoff latency for
throughput when batching.

\subsection{Flash Log}
Cyclone is designed to use only limited amounts of NVM and drains log entries
down into a log placed on a flash SSD - that we term the flashlog. The flashlog
is written out in segments of configurable size (we use 128KB segments). A
segment buffer is prepared in memory (volatile DRAM) using the layout shown in
Figure~\ref{fig:flashlog_page}. We do not allow objects in the flashlog to cross
a 4KB boundary - linking multiple objects together with a special flag encoded
into the size if necessary. We flush log segments out to the log file using
asynchronous direct IO, and therefore we fill one log segment buffer while
keeping IO to another one outstanding to the flash drive. To avoid having to do
a synchronous metadata flush, we preallocate (using the posix\_fallocate call) a
gigabyte worth of zero filled disk pages at the end of a file, at a time.

@@

We move log entries from the head of the NVM log to the flashlog buffers in FIFO
order. THe NVM log entry is only actually removed then the IO for the
corresponding flashlog page is complete. This means that during recovery we can
have the same log entry both in the NVM log and the flashlog, a condition that
can be detected by examining the RAFT related information that is embedded as
part of the logged packet.

@@

The fact that we can immedately absorb updates into an NVM log and only move
them to the flashlog in suitable units for block IO, sidesteps a common problem
with logging to secondary storage - the need to do group commit. Logging systems
often wait to collect a batch of entries to write to secondary storage trading
latency for throughput. This is unnecessary in Cyclone as it effectively makes
use of directly attached NVM with a two level log that removes the need to make
this tradeoff.

@@

\section{Horizontal Scaling}
\label{sec:horizontal}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated CPU core. Further, we assume a concurrent
key value store where a set of CPU cores independently run operations from the
physical logs on the shared memory key-value store data structure, synchronizing
as necessary.

>>>>

\documentclass[10pt, preprint, nonatbib]{sigplanconf}
\usepackage{times,epsfig,endnotes,graphicx, verbatim}
\usepackage{datetime, url, pdfpages, parskip}

@@

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone offers a solution to this problem
by leveraging a small amount of directly attached non-volatile memory (NVM) such
as NVDIMMs to maintain a two level log: the upper level in a small amount of NVM
draining into a larger lower level log maintained on flash. Replicating the
smaller log kept in NVMs transforms the replication problem into a packet
switching problem as the leader sends the exact same packet to all follower
replicas and we solve the problem as such by using the RAFT protocol implemented
as a software multicast module on the Data Plane Development Kit (DPDK). Second,
we leverage the observation that key value store interfaces are commutative for
operations to different keys to scale our two level log horizontally into a
number of physical two level logs - with a hash of the key being used to choose
the physical log. Each log is replicated by an independent instance of RAFT,
thereby leveraging the scalability afforded by classic software packet switching
when using multiple CPU cores. Cyclone is able to replicate millions of small
updates a second using only commodity 10 gigabit ethernet adapters and improves
the performance and availability of RocksDB, a popular persistent key value
store by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

\section{Introduction}
Persistent key value stores are an increasingly important component of
datacenter scale storage services. Key value stores such as
Rocksdb~\cite{rocksdb}, LevelDB~\cite{leveldb} and FloDB~\cite{flodb} represent
large amounts of effort on both the engineering and research fronts. These key
value stores include sophisticated in-memory data structures built around Log
Structured Merge trees~\cite{lsmtree} and are heavily tuned to extract maximum
performance from flash-based solid state drives (SSDs).

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in many internet scale services where
high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write ahead log that if replicated
and kept durable for every appended update provides the necessary high
availability. Unfortunately the write ahead is a performance achilles heel for
these systems - eclipsing much of the work on improving performance of the LSM
component. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked Rocksdb shows the performance of
Rocksdb without the write ahead log. The performance when persisting the write
ahead log without replicating it is shown as the line marked 'rocksdb/WAL'. The
line marked 'rocksdb/3 way rep.' is for simply replicating the log three ways
without persisting it, using RAFT~\cite{raft} running over TCP/IP. Either
persisting every update to the log or replicating it using TCP/IP causes
performance to drop by an order of magnitude (observe the log scale on the
x-axis). It is therefore no suprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store research such as FloDB~\cite{flodb} turns off the write ahead log to be
able to showcase benefits of sophisticated extensions to LSM datastructures.

@@

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone provides the
order of magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the
host - such as in the form of NVDIMMs~\cite{farm} or directly attached
persistent memory~\cite{pmfs}.

Cyclone makes use of a limited amount of attached NVM using two ideas. First, it
leverages the observation that with persistence being a first class property of
a subset of directly addressable DRAM, replicating the same log entry from the
leader to follower replicas is no different from multicast using software packet
switching techniques. It directly leverages this observation to implement a
version of RAFT that replicates the NVM log using classic packet switching
techniques such as separation of control and data planes. Second, Cyclone uses a
two level log structure where it drains entries from the NVM log in IO friendly
batches to a log placed on a flash drive. This allows Cyclone to use only a
limited amount of NVM reasoning that current forms of NVM such as NVDIMMs are
more expensive than DRAM. Our experiments use an NVM footprint of a mere 64 MB
to achieve performance comparable to placing the entire log in NVM. This allows
Cyclone to be used in deployments where the key value store is maintained
entirely in main memory rather than as a multi-level LSM tree structure, with
the log serving as the source for recovery in the event of failure.

We discuss the two level structure focusing on the packet switching components
in Section~\ref{sec:2level}. We then describe how we can horizontally scale this
basic idea across multiple cores by leveraging the commutativity of key value
store APIs across different keys in Section~\ref{sec:horizontal}. We then
demonstrate in the evaluation in Section~\ref{sec:evaluation} that Cyclone can
replicate millions of updates a second using only commodity 10 Gigabit ethernet
adapters thereby closing the performance gap between key value store performance
and write ahead logging performance using only a small amount of directly
attached NVM. A discussion of related work follows before we conclude.

\section{Two Level Log}
\label{sec:2level}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone's construction is predicated on the observation that
in the common case, protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in persistent directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flashlog. We periodically drain entries in batches from the NVM
log to the backing flashlog. An important consequence of this two level approach
is that we need only a limited amounts of NVM with the bulk of the log residing
on flash - an arrangement that maps well to the current relative costs of the
two types of memory. We begin by describing the NVM log and how we implement the
RAFT consensus protocol in the packet switching paradigm around it. We then
describe the flashlog and how we drain entries in batches from the NVM level to
the flash level.

\subsection{NVM Log}
The NVM log is maintained as a circular log of fixed sized pointers (a pointer
log) to memory buffers - each of which holds a client request. This arrangement
is shown in Figure~\ref{fig:nvm_log} that also illustrates the advantage of such
an arragement. Adding a level of indirection in the NVM log allows us to
separate the FIFO ordered circular log being manipulated by the RAFT from
variable sized client being managed by the memory allocator of the packet
switching library we use. Another advantage is that it makes recovery from NVM
easy - appends to the circular pointer log are atomic and we can use the pointer
log to recover allocator state i.e. what pieces of NVM are currently in use by
the log.

@@

We now describe how we implement the RAFT consensus protocol to replicate the
NVM log across a set of replicas. The software packet switching idea we apply is
to separate the control plane from the data plane using the Data Plane
Development Kit (DPDK~\cite{dpdk}) library for direct userspace access to the
network interface.

\subsubsection{Dataplane Separation}
We use RAFT~\cite{raft} as the underlying protocol for log replication. RAFT
replicates a log across a set of replicas. Its operations are leader-based and
RAFT incorporates a leader election algorithm that elects a leader with the most
up to date log. The leader receives log entries from clients, appends them to the
persistent log (NVM log in this case) and multicasts the new entry to follower
replicas. On a response from enough follower replicas (constituting a majority
quorum) the leader replica can then execute the command in the log entry. In
our case this involves accessing the key value store and sending the response
back to the client.

We implement the common case flow in RAFT as a software packet switch using
DPDK. The dataplane now encompasses the Network Interface Card (NIC) and the CPU
caches with the NIC placing packet data directly in the CPU cache via
DDIO~\cite{ddio} - a standard feature of most NICs today. The dataplane for the
RAFT protocol is therefore separated from the control plane. The control plane
runs in a thread pinned to a dedicated CPU core and takes care of the protocol
state machine: checking for correct term, log index and that the replica
receiving the packet from the client is indeed the leader. On the follower side
it is responsible for checking that the leader is in the correct term before
accepting the packet by appending it to the NVM log and initiating a
response. In addition, the control plane is responsible for uncommon case
activities such as dealing with timeouts and leader elections.

The pseudocode in Figure~\ref{fig:control_plane} describes part of the control
plane in Cyclone organized as event handlers triggered on receiving a packet at
the leader. We focus on one event for illustration (and brevity): the event
where a request is received from a client. A key concern when we built Cyclone
was adhering as far as possible to the software packet switching principle of
separating the control plane from the dataplane. This means that the CPU should
avoid examining every byte of the packet as far as possible.

The sequence of operations when a packet is received from the client needs
careful design for efficient software packet switching. To illustrate how this
is achieved, Figure~\ref{fig:packet_layout} shows how Cyclone manipulates packet
layouts across the two steps of prepending a RAFT header and transmitting to
follower replicas. DPDK describes packets using an ``mbuf'' data
structure. Roughly speaking, an ``mbuf'' consists of a flat array of bytes
actually containing the packet and a fixed size piece of external metadata that
describes various aspects of the packet, most crucially a pointer to the start
and end of the packet in the byte array. DPDK's userspace drivers receive
packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We strip off the existing
network headers in the packet and prepend RAFT related information specific to
each log entry in the headroom by shifting the start pointer appropriately.
These operations are standard enough for software packet switches that DPDK
provides convenient library calls for it. For the final step, we need to prepare
the packet for transmission to the various follower replicas. To do this we
prepare a different packet containing an ethernet header for \emph{each}
targeted replica and ``chain'' the data packet to each of these headers. Each
header is then separately handed off to the driver for transmission via the NIC,
carrying the data packet with it by association.

\begin{figure}
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Cyclone Control Plane Fragment}
\label{fig:control_plane}
\end{figure}

@@

We note that by discrding stream oriented protocols such as TCP, the network
layer is not longer reliable or ordered. This is not a problem for RAFT that is
designed to tolerate network failures resulting in drops or reordering of
network packets. However, this rules out consensus protocols such as Zookeeper
Atomic Broadcast (ZAB~\cite{zab}) that depend on a reliable stream-oriented
network layer to function correctly.

@@

We note that batching in Cyclone does not involve a latency-throughput
tradeoff like in many other systems~\cite{ix-dataplane}. The batch receive call
we use in DPDK returns immediately with whatever number of packets is available,
including zero. We always flush the transmit buffer after every call to DPDK to
transmit packets to replicas. Therefore, we never tradeoff latency for
throughput when batching.

\subsection{Flash Log}
Cyclone is designed to use only limited amounts of NVM and drains log entries
down into a log placed on a flash SSD - that we term the flashlog. The flashlog
is written out in segments of configurable size (we use 128KB segments). A
segment buffer is prepared in memory (volatile DRAM) using the layout shown in
Figure~\ref{fig:flashlog_page}. We do not allow objects in the flashlog to cross
a 4KB boundary - linking multiple objects together with a special flag encoded
into the size if necessary. We flush log segments out to the log file using
asynchronous direct IO, and therefore we fill one log segment buffer while
keeping IO to another one outstanding to the flash drive. To avoid having to do
a synchronous metadata flush, we preallocate (using the posix\_fallocate call) a
gigabyte worth of zero filled disk pages at the end of the log file whenever we
hit its end.

@@

We move log entries from the head of the NVM log to the flashlog buffers in FIFO
order. The NVM log entry is only actually removed when the IO for the
corresponding flashlog page is complete. This means that during recovery we can
have the same log entry both in the NVM log and the flashlog, a condition that
can be detected by examining the RAFT related information that is embedded as
part of the logged packet.

@@

The fact that we can immedately absorb updates into an NVM log and only move
them to the flashlog in suitable units for block IO sidesteps a common problem
with logging to secondary storage - the need to do group commit. Logging systems
often wait to collect a batch of entries to write to secondary storage trading
latency for throughput. This is unnecessary in Cyclone as the use of directly
attached NVM with a two level log removes the need to make this tradeoff.

@@

\section{Horizontal Scaling}
\label{sec:horizontal}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section, we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated CPU core. Further, we assume a concurrent
key value store where a set of CPU cores independently run operations from the
physical logs on the shared memory key-value store data structure, synchronizing
among themselves as necessary.
